---
title: "Package Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Functions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

To install the package, please use: 

```
library(devtools)
install_github("doomlab/ViSe")
```

This installation will give you the latest version of the package, regardless of status on CRAN. 

```{r setup}
library(ViSe)
library(flextable)
library(cowplot)
library(ggplot2)
```

## Example

To demonstrate the use of visual sensitivity analysis, we use a study of the effect of child maltreatment on the extent of mental health problems in terms of internalising and externalising behaviour. The study of Kisely and colleagues (2018) was based on a general population sample in Brisbane, Australia, and compared 3554 mother-child pairs without 'substantiated child maltreatment' to, for example, 73 pairs with child neglect. Note that the results vary across different types of maltreatment assessed, we choose child neglect because its results (a smaller but still considerable l after adjustment) are particularly suited to illustrating sensitivity analysis. Maltreatment was assessed 'by linkage to state child protection agency data'. Internalising and externalising behaviours were measured using the Youth Self-Report (YSR) scale at around the age of 21. The study reports unadjusted mean differences and mean differences adjusted for ‘gender, parental ethnicity, maternal age, family income, maternal relationship status, maternal education, youth income level, youth education level, youth marital status’ (e.g., likelyprobably based on ordinary least squares regression, but the paper does not specify that). To obtain the estimates and two-tailed 95% confidence intervals on the d scale, we divided the reported unadjusted and adjusted means for internalising and externalising by the respective standard deviations of the total sample (see Table \@ref(tab:table-1)) with the `calculate_d` function.

See supplemental document at https://static.cambridge.org/content/id/urn:cambridge.org:id:article:S0007125018002076/resource/name/S0007125018002076sup001.docx 

## Calculate effect size 

### Unadjusted scores

```{r}
# internalising
internal_unadj <- 
  calculate_d(m1 = 14.37, # neglect mean
              sd1 = 10.716, # neglect sd
              n1 = 71, # neglect n
              m2 = 10.69, # none mean
              sd2 = 8.219, # none sd
              n2 = 3653, # none n
              a = .05, # alpha/confidence interval
              lower = TRUE) # lower or upper bound

internal_unadj$d # effect size 
internal_unadj$dlow # non-directional non-central lower 
internal_unadj$dlow_central # non-directional central lower 
internal_unadj$done_low # one-tailed non-central lower
internal_unadj$done_low_central # one-tailed central lower

# externalising
external_unadj <- 
  calculate_d(m1 = 13.14, sd1 = 8.066, n1 = 71, 
            m2 = 9.42, sd2 = 6.764, n2 = 3653, 
            a = .05, lower = TRUE)

external_unadj$d # effect size 
external_unadj$dlow # non-directional non-central lower 
external_unadj$dlow_central # non-directional central lower 
external_unadj$done_low # one-tailed non-central lower
external_unadj$done_low_central # one-tailed central lower
```

Note, you can also calculate directly from a t-test output in *R*, a *t*-value, or the dataframe of scores. 

### Adjusted scores

The adjusted scores are not displayed in raw score format, but rather mean difference scores. We can again use calculate_d in a different way:

```{r}
# internalising
m_diff <- 2.73
lower <- 0.77
# ci is m - t_alpha*SE = lower
SE <- (m_diff - lower)/qt(.025, # .05 / 2
                    df = (71+3653 - 2), # n - 1 + n - 1
                    lower.tail = FALSE) 
SE

# confirm correct
# p value is listed as .006
pt(q = m_diff / SE, # t value
   df = (71+3653 - 2), # df 
   lower.tail = FALSE)*2 # upper tail * 2 for two tails 

internal_adj <- 
  calculate_d(t = m_diff/SE, # t-score
            n1 = 71, # neglect n
            n2 = 3653, # none n
            a = .05,  # alpha/confidence interval
            lower = TRUE) # lower or upper bound

internal_adj$d # effect size 
internal_adj$dlow # non-directional non-central lower 
internal_adj$dlow_central # non-directional central lower 
internal_adj$done_low # one-tailed non-central lower
internal_adj$done_low_central # one-tailed central lower

# externalising
m_diff <- 3.10
lower <- 1.49
# ci is m - t_alpha*SE = lower
SE <- (m_diff - lower)/qt(.025, # .05 / 2
                    df = (71+3653 - 2), # n - 1 + n - 1
                    lower.tail = FALSE) 
SE

# confirm correct
# p value is listed as < .001
pt(q = m_diff / SE, # t value
   df = (71+3653 - 2), # df 
   lower.tail = FALSE)*2 # upper tail * 2 for two tails 

external_adj <- 
  calculate_d(t = m_diff/SE, # t-score
            n1 = 71, # neglect n
            n2 = 3653, # none n
            a = .05,  # alpha/confidence interval
            lower = TRUE) # lower or upper bound

external_adj$d # effect size 
external_adj$dlow # non-directional non-central lower 
external_adj$dlow_central # non-directional central lower 
external_adj$done_low # one-tailed non-central lower
external_adj$done_low_central # one-tailed central lower
```

## Visualisation of c

```{r}
plots_1 <- visualize_c(dlow = internal_unadj$dlow_central)$graph + 
  ggtitle("Internal Unadjusted") + 
  annotate("text", label = paste0("l = ", round(internal_unadj$dlow_central, 2)), 
           x = -1.5, y = -.5)
plots_2 <- visualize_c(dlow = external_unadj$dlow_central)$graph + 
  ggtitle("External Unadjusted") + 
  annotate("text", label = paste0("l = ", round(external_unadj$dlow_central, 2)), 
           x = -1.5, y = -.5)
plots_3 <- visualize_c(dlow = internal_adj$dlow_central)$graph + 
  ggtitle("Internal Adjusted") + 
    annotate("text", label = paste0("l = ", round(internal_adj$dlow_central, 2)), 
           x = -1.5, y = -.5)
plots_4 <- visualize_c(dlow = external_adj$dlow_central)$graph + 
  ggtitle("External Adjusted") + 
  annotate("text", label = paste0("l = ", round(external_adj$dlow_central, 2)), 
         x = -1.5, y = -.5)

plot_grid(plots_1, plots_2, 
          plots_3, plots_4, 
          nrow = 2)

ggsave(filename = "visualize_c.png")
```
## Calculation of adjusted values

```{r}
adjusted_coef(effect_xz = .2, # effect of x and z
    effect_uxz = .4, # effect of x, u, and z
    effect_d = .12) # effect size of difference score between groups 
```

## Calculation between effect sizes

```{r}
d_to_f2(.25)
d_to_nnt(.25)
d_to_r(.25)
```

